String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-08-30 2023-1-19/sections/abstract.tex
some updates
add files

%!TEX root = ../main.tex

Regular expressions (regex for short) and string-length function are widely used in string-manipulating programs. Counting is a frequently used feature in regexes that counts the number of matchings of sub-patterns. 
The state-of-the-art string solvers are incapable of solving string constraints with regex-counting and string-length eﬀiciently, especially when the counting and length bounds are large. 
In this work, we propose an automata-theoretic approach for solving such a class of string constraints. 
The main idea is to model the counting operators symbolically by registers in automata instead of unfolding them explicitly, thus alleviating the state explosion problem. Moreover, the string-length function is modeled by a register as well. 
As a result, the satisfiability of string constraints with regex-counting and string-length is reduced to the satisfiability of linear integer arithmetic, which the off-the-shelf SMT solvers can then solve. 
To improve the performance further, we propose techniques to reduce automata sizes. We implement the algorithms and validate our approach on 48,843 benchmark instances. 
The experimental results show that our approach can solve more instances than the other solvers, at a comparable or faster speed, especially when the counting and length bounds are large.

>>>>

Regular expressions (regex for short) and string-length function are widely used in string-manipulating programs. Counting is a frequently used feature in regexes that counts the number of matchings of sub-patterns. 
The state-of-the-art string solvers are incapable of solving string constraints with regex-counting and string-length eﬀiciently, especially when the counting and length bounds are large. 
In this work, we propose an automata-theoretic approach for solving such class of string constraints. 
The main idea is to symbolically model the counting operators by registers in automata instead of unfolding them explicitly, thus alleviating the state explosion problem. Moreover, the string-length function is modeled by a register as well. 
As a result, the satisfiability of string constraints with regex-counting and string-length is reduced to the satisfiability of linear integer arithmetic, which the off-the-shelf SMT solvers can then solve. 
To improve the performance further, we propose simplification techniques to reduce automata sizes. We implement the algorithms and validate our approach on 48,843 benchmark instances. 
The experimental results show that our approach can solve more instances than the state-of-the-art solvers, at a comparable or faster speed, especially when the counting and length bounds are large.

%\keywords{Rgular Expressions, Counting, Length, Automata, Registers, Linear Integer Arithmetic}
%\end{document}

### 2023-08-30 2023-1-19/sections/implementation.tex
some updates
add files

%!TEX root = ../main.tex
%\documentclass{standalone}
%\begin{document}

We implement the algorithms in Section~\ref{sec:algorithm} on top of OSTRICH~\cite{Ostrich,atva2020}, resulting in a version of OSTRICH called $\ostrichrecl$.
Then we do experiments on two benchmark suites to evaluate the performance of $\ostrichrecl$, and carry out detailed experiments, including the evaluation of size-reduction techniques and the experiment of large-counting benchmark. In the sequel, we first describe the two benchmark suites, then present the settings and results of the experiments.

@@

\vspace{-3mm}
\subsection{Benchmark Suites and Experiment Setup}\label{sec:bench}
\vspace{-2mm}

Our experiments utilize two benchmark suites, namely, RegCoL and AutomatArk. In total, there are 48,843 instances in the two benchmark suites. All the benchmark instances are in SMTLIB2 format.

\medskip
\noindent
\emph{RegCoL benchmark suite.} There are 40,628 RECL constraints in the RegCoL suite. These constraints are generated by extracting regular expressions with counting operators from the open source regex library \cite{regex_lingua_franca,redos_lenka} and constructing for each regular expression $e$ a RECL constraint of the form
$x \in e \wedge x \in \overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*} \wedge |x| > 10,$
where $\overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*}$ is a regular expression that sanitizes all occurrence of special characters $<$, $>$, $'$, $''$, or $\&$. The artificial construction is meaningful because the special characters are offensive. For instance, they are widely used in client-side attacks like XSS\cite{malware_detection_3_kudzu}\cite{CCH+18}.
%We would like to remark that there are about 500,000 real-world regular expressions in \cite{regex_lingua_franca,redos_lenka} and regular expressions with counting operators occupy about 8\% of them.

\medskip
\noindent
\emph{AutomatArk benchmark suite.}
The AutomatArk benchmark suite is adapted from the AutomatArk suite in \cite{z3str3re} by filtering out the string constraints that contain no occurrences of counting operators and adding the length constraint $|x| > 10$ (where $x$ is a string variable). There are 8,215 instances in the AutomatArk suite here.
Note that the original AutomatArk benchmark suite in \cite{z3str3re} includes 19,979 instances, conjunctions of regular membership queries generated out of regular expressions from \cite{automatark}.
%Only 8,751 contain counting operators out of 19,979 instances. 
Moreover, only 5\% of them include non-register-representable regexes among the 48,843 instances.
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%

@@

\vspace{-3mm}
\subsection{Overall Experiment Results}
\vspace{-2mm}

We evaluate the performance of $\ostrichrecl$ against the state-of-the-art string constraint solvers, including CVC5
%(version 1.0.5) 
\cite{cvc5}, Z3seq \cite{z3seq}, Z3str3
%(github commit 59e9c87) 
\cite{z3str3}, Z3str3RE \cite{z3str3re}, and OSTRICH
%(github commit 8297d8d) 
\cite{ostrich2023}, on RegCoL and AutomatArk benchmark suites.
% Z3-trau \cite{z3trau} is not evaluated because it does not support \verb|re.diff|, the language difference operator of regular expressions.  
%The experiments are designed to answer the following two research questions. 
%\begin{description}
%\item[Q1] Does OSTRICH$^{\rm RECL}$ solve RECL constraints more efficiently than the state-of-the-art string constraint solvers ?
%
%\item[Q2] Do the size-reduction and under-approximation techniques proposed in Section~\ref{sec:algorithm} indeed improve the performance ?
%\end{description}
\vspace{-5mm}
\begin{table}
  \centering
  \begin{subtable}{0.513\textwidth}
      \centering
      \import{tables}{table_regcol.tex}
      \caption{Overall experiment results,\\ with timeout = 60 seconds.}
      \label{tab:results_regcol}
  \end{subtable}
  \begin{subtable}{0.477\textwidth}
      \centering
      \import{tables}{table_simp.tex}
      \caption{Simplification technique evaluation, with timeout = 60 seconds.}
      \label{tab:results_simp}
  \end{subtable}
  \vspace{-6mm}
  \caption{Experiments on RegCoL and AutomatArk benchmark suites}
\end{table}
\vspace{-9mm}
% \begin{table}[ht]
%   \vspace{-5mm}
%   \begin{center}
%     \import{tables}{table_regcol.tex}
%   \end{center}
%   \caption{Overall Experiment results, with timeout = 60 seconds}
%   \label{tab:results_regcol}
%   \vspace{-9mm}
% \end{table}

The overall experiment results are in Table~\ref{tab:results_regcol}. Note that we take the result of CVC5 as the ground truth\footnote{At first we used the majority vote between all results of the solvers as the ground truth, but solvers of z3 family output some wrong results(verified by us manually), causing the ground truth to be incorrect.}. We can see that $\ostrichrecl$ solves almost all 48,843 instances, except 182 of them, that is, it solves \textbf{48,662} instances correctly. The number is %3,994/2,272/24,577/12,832/1,091/7,092 
\textbf{3,908/1,111/12,838/2,306/2,396 more} than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
%      
%A soundness error is reported if the result of a solver for an instance is different from CVC5. 
% their results on some instances are inconsistent with the results of the other three solvers, i.e. CVC5, OSTRICH, and OSTRICH$^{\rm RECL}$.
%
Moreover, our solver $\ostrichrecl$ is the second fastest solver, and its average time on each instance is close to the fastest solver Z3str3RE (\textbf{1.93}/1.62 s).
\vspace{-3mm}
\subsection{Evaluation of Simplification Techniques}
\vspace{-2mm}

We evaluate the simplification techniques by comparing the performance of $\ostrichrecl$ with and without the simplification techniques. Moreover, we also encode the $\cefadec$ problem to an infinite state transition system and solve it using \textsc{nuXmv}\cite{nuxmv}. 
% \begin{table}[ht]
%   \vspace{-5mm}
%   \begin{center}
%     \import{tables}{table_simp.tex}
%   \end{center}
%   \caption{Simplification technique evaluation, with timeout = 60 seconds}
%   \label{tab:results_simp}
%   \vspace{-9mm}
% \end{table}

The evaluation results of simplification techniques are in Table~\ref*{tab:results_simp}. Compared to $\ostrichrecl$ without simplification techniques, $\ostrichrecl$ with simplification techniques solves \textbf{1,503 more} instances and is \textbf{2.21$\times$} faster. Moreover, compared with the techniques using \textsc{nuxmv}, $\ostrichrecl$ with simplification techniques solves \textbf{1,798 more} instances and is \textbf{3.13$\times$} faster.

\vspace{-3mm}
\subsection{Large-Counting Experiment Results}
The distribution of upper bounds of counting in RegCoL and AutomatArk suites is shown in Fig~\ref{fig:count_distri}, whose horizontal axis is the value of the upper bound and vertical axis is the number of occurrences. The upper bound is the number $n$ of regex $e^{\{m,n\}}, e^{\{n,n\}}$ and $e^{\{n,\infty\}}$ (which is equivalent to $e^{\{n,n\}}e^*$). Although most upper bounds of counting operators are less than 50, about two thousand are still greater than 50. We extract instances with upper bounds greater than 50 and generate a \emph{`large-counting'} benchmark suite (1,969 instances in total). Moreover, We change the length constraint of each instance to $|x| > 200$ (which is $|x| > 10$ in the overall experiment), to make the suite more balanced.
\vspace{-5mm}
\begin{figure}[ht]
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \includegraphics[width=1\textwidth]{counting_distribution.png}  
    \caption{Distribution of counting operators in two benchmark suites.}  
    \label{fig:count_distri}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \import{tables}{table_large_count.tex}
    \vspace{4.5mm}
    \caption{Large-counting experiment, \\with timeout = 60 seconds.}
    \label{fig:table_large_count}
  \end{subfigure}
  \vspace{-2mm}
  \caption{Analysis of dataset and large-counting experiment.}
\end{figure}
\vspace{-5mm}

@@

  To answer Q2, we compare OSTRICH$^{\rm RECL}$ with its three variants, namely, OSTRICH$^{\rm RECL}$ where the size-reduction, under-approximation, or both are removed.
  For convenience, let us use OSTRICH$^{\rm RECL^-}$ to denote the variant of OSTRICH$^{\rm RECL}$ that both size-reduction and under-approximation techniques are removed,
  OSTRICH$^{\rm RECL^-}_{+\rm SR}$ (resp. OSTRICH$^{\rm RECL^-}_{+\rm UA}$) to denote the variant that is obtained by adding the size-reduction (resp. under-approximation) techniques to OSTRICH$^{\rm RECL^-}$.
  %OSTRICH$^{\rm RECL}_{\rm -SR}$, OSTRICH$^{\rm RECL}_{\rm -UA}$, OSTRICH$^{\rm RECL}_{\rm -SRUA}$, which remove size-reduction, under-approximation, and both from OSTRICH$^{\rm RECL}$ respectively. 
  The experiment results can be found in Table~\ref{tab:results_heuristics}.
  From Table~\ref{tab:results_heuristics}, we can see that adding size-reduction techniques help to solve 1,563 more instances and reduce the average time per instance for 46.4\% (see the OSTRICH$^{\rm RECL^-}_{\rm +SR}$ column), while adding under-approximation techniques help to solve 1,476 more instances and reduce the average time per instance for 42.7\% (see the OSTRICH$^{\rm RECL^-}_{\rm +UA}$ column), moreover, adding both help to solve 1,579 more instances and reduce the average time per instance for 46.5\% (see the OSTRICH$^{\rm RECL}$ column). Note that adding under-approximation to OSTRICH$^{\rm RECL^-}_{\rm +SR}$ helps to solve 19 more sat instances while reducing the number of solved unsat instances by 3 since under approximation techniques do not help to solve unsat instances.
}
%%%%%%%%%%%%%%%%%% Q2 removed %%%%%%%%%%%%
%%%%%%%%%%%%%%%%%% Q2 removed %%%%%%%%%%%%
%Moreover, adding these techniques also decrease the average time spent in solving each instance considerably.

>>>>

%!TEX root = ../main.tex
%\documentclass{standalone}
%\begin{document}
In this section, we list our experimental results. We first present the overall experiment results on RegCoL and AutomatArk benchmark suites. Then we evaluate the simplification techniques and large-counting experiment results.

@@

\vspace{-2mm}
\subsection{Benchmark Suites and Experiment Setup}\label{sec:bench}
\vspace{-1mm}

Our experiments utilize two benchmark suites, namely, \emph{RegCoL} and \emph{AutomatArk}. There are 48,843 instances in total, and all benchmark instances are in SMTLIB2 format.

\medskip
\noindent
\emph{RegCoL benchmark suite.} There are 40,628 RECL instances in the RegCoL suite. These instances are generated by extracting regexes with counting operators from the open source regex library \cite{regex_lingua_franca,redos_lenka} and manually constructing a RECL constraint $x \in e \wedge x \in e_{sani} \wedge |x| > 10$ for each regex $e$,
where $e_{sani} \equiv \overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*}$ is a regular expression that sanitizes all occurrence of special characters $<$, $>$, $'$, $''$, or $\&$. The artificial construction is meaningful because the special characters are offensive. For instance, they are widely used in client-side attacks like XSS\cite{malware_detection_3_kudzu,CCH_18}.
%We would like to remark that there are about 500,000 real-world regular expressions in \cite{regex_lingua_franca,redos_lenka} and regular expressions with counting operators occupy about 8\% of them. 
\vspace{-0.1mm}

\medskip
\noindent
\emph{AutomatArk benchmark suite.}
The AutomatArk benchmark suite is adapted from the AutomatArk suite \cite{z3str3re} by picking out the string constraints containing counting operators. We add length constraint $|x| > 10$ for each string variable $x$ because there is no original length constraints. There are 8,215 instances in the AutomatArk suite.
Note that the original AutomatArk benchmark suite \cite{z3str3re} includes 19,979 instances, conjunctions of regular membership queries generated out of regular expressions from \cite{automatark}.
%Only 8,751 contain counting operators out of 19,979 instances. 
Moreover, only 5\% of regexes among the 48,843 instances are non-register-representable.
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%

@@

\vspace{-2mm}
\subsection{Overall Experiment Results}
\vspace{-1mm}

We evaluate the performance of $\ostrichrecl$ against the state-of-the-art string constraint solvers, including CVC5
%(version 1.0.5) 
\cite{cvc5}, Z3seq \cite{z3seq}, Z3str3
%(github commit 59e9c87) 
\cite{z3str3}, Z3str3RE \cite{z3str3re}, and OSTRICH
%(github commit 8297d8d) 
\cite{ostrich2023}, on RegCoL and AutomatArk benchmark suites.
% Z3-trau \cite{z3trau} is not evaluated because it does not support \verb|re.diff|, the language difference operator of regular expressions.  
%The experiments are designed to answer the following two research questions. 
%\begin{description}
%\item[Q1] Does OSTRICH$^{\rm RECL}$ solve RECL constraints more efficiently than the state-of-the-art string constraint solvers ?
%
%\item[Q2] Do the simplification and under-approximation techniques proposed in Section~\ref{sec:algorithm} indeed improve the performance ?
%\end{description}
\vspace{-5mm}
\begin{table}
  \centering
  \begin{subtable}{0.513\textwidth}
      \centering
      \import{tables}{table_regcol.tex}
      \caption{Overall experiment results,\\ with timeout = 60 seconds.}
      \label{tab:results_regcol}
  \end{subtable}
  \begin{subtable}{0.477\textwidth}
      \centering
      \import{tables}{table_simp.tex}
      \caption{Simplification techniques evaluation, with timeout = 60 seconds.}
      \label{tab:results_simp}
  \end{subtable}
  \vspace{-6mm}
  \caption{Experiments on RegCoL and AutomatArk benchmark suites}
\end{table}
\vspace{-9mm}
% \begin{table}[ht]
%   \vspace{-5mm}
%   \begin{center}
%     \import{tables}{table_regcol.tex}
%   \end{center}
%   \caption{Overall Experiment results, with timeout = 60 seconds}
%   \label{tab:results_regcol}
%   \vspace{-9mm}
% \end{table}

The overall experiment results are in Table~\ref{tab:results_regcol}. Note that we take the result of CVC5 as the ground truth\footnote{At first we used the majority vote between all results of the solvers as the ground truth, but solvers of z3 family output some wrong results(verified by us manually), causing the ground truth to be incorrect.}. We can see that $\ostrichrecl$ solves almost all 48,843 instances, except 182 of them, that is, it solves \textbf{48,662} instances correctly. The number is %3,994/2,272/24,577/12,832/1,091/7,092 
\textbf{3,908/1,111/12,838/2,306/2,396 more} than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
%      
%A soundness error is reported if the result of a solver for an instance is different from CVC5. 
% their results on some instances are inconsistent with the results of the other three solvers, i.e. CVC5, OSTRICH, and OSTRICH$^{\rm RECL}$.
%
Moreover, $\ostrichrecl$ is the second fastest solver, whose average time on each instance is close to the fastest solver Z3str3RE (\textbf{1.93 s}/1.62 s).
\vspace{-2mm}
\subsection{Evaluation of Simplification Techniques}
\vspace{-1mm}

We evaluate the simplification techniques by comparing the performance of $\ostrichrecl$ with and without the simplification techniques. Moreover, we encode the $\cefadec$ problem to an infinite state transition system and use \textsc{nuXmv}\cite{nuxmv} to solve it. We also compare the simplification techniques to the \textsc{nuXmv}-based techniques.
% \begin{table}[ht]
%   \vspace{-5mm}
%   \begin{center}
%     \import{tables}{table_simp.tex}
%   \end{center}
%   \caption{Simplification technique evaluation, with timeout = 60 seconds}
%   \label{tab:results_simp}
%   \vspace{-9mm}
% \end{table}

The evaluation results of simplification techniques are shown in Table~\ref*{tab:results_simp}. $\ostrichrecl_{\rm -SIMP}$ is the tool without simplification techniques. $\ostrichrecl_{\rm NUXMV}$ is the tool using \textsc{nuXmv}-based techniques. We can see that our solver solves \textbf{1,503 more} instances and is \textbf{2.21$\times$} faster than it without simplification techniques. Moreover, compared to the \textsc{nuXmv}-based techniques tool, our solver solves \textbf{1,798 more} instances and is \textbf{3.13$\times$} faster.

\vspace{-2mm}
\subsection{Large-Counting Experiment Results}
\vspace{-1mm}
The distribution of upper bounds of counting in RegCoL and AutomatArk suites is shown in Fig~\ref{fig:count_distri}, whose horizontal axis is the value of the upper bound and vertical axis is the number of occurrences. The upper bound is the number $n$ of regex $e^{\{m,n\}}, e^{\{n,n\}}$ and $e^{\{n,\infty\}}$ (which is equivalent to $e^{\{n,n\}}e^*$). Although most upper bounds are small, about two thousand of them are still greater than 50. We extract these instances containing regex with upper bound greater than 50 and generate a \emph{`large-counting'} benchmark suite (1,969 instances in total). Moreover, We change the length constraint of each instance to $|x| > 200$ (which is $|x| > 10$ in the overall experiment), to make the suite harder to solve.
\vspace{-5mm}
\begin{figure}[ht]
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \includegraphics[width=1\textwidth]{counting_distribution.png}  
    \caption{Distribution of counting operators in two benchmark suites.}  
    \label{fig:count_distri}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \import{tables}{table_large_count.tex}
    \vspace{4.5mm}
    \caption{Large-counting experiment, \\with timeout = 60 seconds.}
    \label{fig:table_large_count}
  \end{subfigure}
  \vspace{-2mm}
  \caption{Analysis of dataset and large-counting experiment.}
\end{figure}
\vspace{-5mm}

@@

  To answer Q2, we compare OSTRICH$^{\rm RECL}$ with its three variants, namely, OSTRICH$^{\rm RECL}$ where the simplification, under-approximation, or both are removed.
  For convenience, let us use OSTRICH$^{\rm RECL^-}$ to denote the variant of OSTRICH$^{\rm RECL}$ that both simplification and under-approximation techniques are removed,
  OSTRICH$^{\rm RECL^-}_{+\rm SR}$ (resp. OSTRICH$^{\rm RECL^-}_{+\rm UA}$) to denote the variant that is obtained by adding the simplification (resp. under-approximation) techniques to OSTRICH$^{\rm RECL^-}$.
  %OSTRICH$^{\rm RECL}_{\rm -SR}$, OSTRICH$^{\rm RECL}_{\rm -UA}$, OSTRICH$^{\rm RECL}_{\rm -SRUA}$, which remove simplification, under-approximation, and both from OSTRICH$^{\rm RECL}$ respectively. 
  The experiment results can be found in Table~\ref{tab:results_heuristics}.
  From Table~\ref{tab:results_heuristics}, we can see that adding simplification techniques help to solve 1,563 more instances and reduce the average time per instance for 46.4\% (see the OSTRICH$^{\rm RECL^-}_{\rm +SR}$ column), while adding under-approximation techniques help to solve 1,476 more instances and reduce the average time per instance for 42.7\% (see the OSTRICH$^{\rm RECL^-}_{\rm +UA}$ column), moreover, adding both help to solve 1,579 more instances and reduce the average time per instance for 46.5\% (see the OSTRICH$^{\rm RECL}$ column). Note that adding under-approximation to OSTRICH$^{\rm RECL^-}_{\rm +SR}$ helps to solve 19 more sat instances while reducing the number of solved unsat instances by 3 since under approximation techniques do not help to solve unsat instances.
}
%%%%%%%%%%%%%%%%%% Q2 removed %%%%%%%%%%%%
%%%%%%%%%%%%%%%%%% Q2 removed %%%%%%%%%%%%
%Moreover, adding these techniques also decrease the average time spent in solving each instance considerably.

### 2023-08-30 2023-1-19/sections/introduction.tex
add files

The string data type plays a crucial role in modern programming languages, such as JavaScript, Python, Java, and PHP. 
String manipulations are error-prone and could even give rise to severe security vulnerabilities (e.g. \ cross-site scripting, aka XSS). 
One powerful method for identifying such bugs is \emph{symbolic execution} (possibly in combination with dynamic analysis). It analyses symbolic paths in a program by viewing them as constraints checked by constraint solvers. 
%
Symbolic execution of string manipulating programs has motivated the highly active research area of \emph{string constraint solving}, resulting in the development of numerous string solvers in the last decade, e.g.,
Z3seq~\cite{z3seq}, CVC4/5~\cite{cvc4}, Z3str/2/3/4~\cite{Z3-str,Z3-str2,Z3-str3,BerzishMurphy2021}, Z3str3RE~\cite{BD+23}, 
Z3-Trau~\cite{Z3-trau}, OSTRICH~\cite{CHL+19}, Slent~\cite{WC+18}, among many others.

@@

\medskip
\noindent 
\emph{Contribution.} In this work, we focus on the class of string constraints with regular membership and string-length function, where the counting operators may occur (called RECL for brevity). We make the following contributions: 
\vspace{-0.5mm}
\begin{itemize}
  \item We propose an automata-theoretical approach for solving RECL constraints. 
  Our main idea is to represent the counting operators by registers in cost-enriched finite automata (CEFA, see Section~\ref{sec:automaton} for the definition), instead of unfolding them explicitly. The satisfiability of RECL constraints is reduced to the nonemptiness problem of CEFA w.r.t. a linear integer arithmetic (LIA), which is then transformed to the satisfiability of existential LIA formula tackled by the off-the-shelf SMT solver.
  % Moreover, the string-length function can also be modeled by register in CEFA. Then the satisfiability of RECL constraints is reduced to the nonemptiness problem of CEFA with respect to a linear integer arithmetic (LIA) formula whose free variables are registers. According to the results from~\cite{atva2020}, an LIA formula can be computed to represent the potential values of registers in CEFA. 
  % Thus, the nonemptiness of CEFA w.r.t. LIA formulas can be reduced to the satisfiability of LIA formulas, which is then tackled by off-the-shelf SMT solvers.
  \item We propose size-reduction techniques to reduce the state space and the number of transitions in CEFA, to achieve better performance. It is beneficial since the sizes of LIA formulas depend much on the sizes of CEFA.
  \item Combined with the size-reduction techniques mentioned above, the register representation of regex-counting and string-length in CEFA entails an efficient algorithm for solving RECL constraints. We implement the algorithm on top of OSTRICH, resulting in a practical string solver called $\ostrichrecl$. 
  \item Finally, we utilize a collection of benchmark suites comprising 48,843 instances in total to evaluate the performance of our tool. The result shows that $\ostrichrecl$ solves the most instances among all the solvers, with comparable speed (See Table~\ref{tab:results_regcol}). In addition, we carry out some detailed experiments, including the evaluation of size-reduction techniques and the experiment of large-counting benchmarks. The result shows that the size-reduction techniques are effective in practice (See Table~\ref{tab:results_simp}), and $\ostrichrecl$ performs much better than other solvers on the large-counting benchmark (See Fig~\ref{fig:table_large_count}).
  % The experiment results show that $\ostrichrecl$ can solve 48,698 out of 48,843 instances, while Z3str3RE, the state-of-the-art solver on RECL constraints, solves only 47,607 instances (where the timeout period is set as 60 seconds). Moreover, the speed of $\ostrichrecl$ is comparable to that of Z3str3RE. In comparison, CVC5, Z3seq, Z3-Trau, and Z3str3 solve only 44,704, 46,426, 24,121, and 35,866 instances respectively. (See Table~\ref{tab:results_regcol}-\ref{tab:results_summary} for more details.)
\end{itemize}
\vspace{-0.5mm}
% We propose an automata-theoretical approach for solving RECL constraints. Our main idea is to represent the counting operators by registers in automata (as much as possible) instead of unfolding them explicitly. Our approach is built upon cost-enriched finite automata (CEFA), proposed in~\cite{atva2020} to solve string constraints involving the integer data type. Moreover, the string-length function can also be modeled by register in CEFA. Then the satisfiability of RECL constraints is reduced to the nonemptiness problem of CEFA with respect to a linear integer arithmetic (LIA) formula whose free variables are registers of CEFA.

@@

% Combined with the size-reduction techniques mentioned above, the register representation of regex-counting and string-length in CEFA entails an efficient algorithm for solving RECL constraints. We implement the algorithm on top of OSTRICH, resulting in a new version of OSTRICH called $\ostrichrecl$ for convenience. 
% Furthermore, we utilize a collection of benchmark suites comprising 48,843 RECL constraints in total to evaluate the performance of $\ostrichrecl$. It includes a widely-used benchmark suite for string constraints and benchmark suites comprising constraints generated from regular expressions in various open source regex libraries. 
% %
% %Note that in the benchmark suites, we only consider the regexes that contain explicit counting, that is, regexes that contain sub-expressions of the form $e^{\{n, m\}}$, and ignore those that contain only implicit counting operators, that is, $e^*$ or $e^+$, since the state-of-the-art string solvers suffer mainly from the explicit counting operators. 
% %
% The experiment results show that $\ostrichrecl$ can solve 48,698 out of 48,843 instances, while Z3str3RE, the state-of-the-art best solver on RECL constraints, solves only 47,607 instances (where the timeout period is set as 60 seconds). Moreover, the speed of $\ostrichrecl$ is comparable to that of Z3str3RE. In comparison, CVC5, Z3seq, Z3-Trau, and Z3str3 solve only 44,704, 46,426, 24,121, and 35,866 instances respectively. (See Table~\ref{tab:results_regcol}-\ref{tab:results_summary} for more details.)

>>>>

The string data type plays a crucial role in modern programming languages such as JavaScript, Python, Java, and PHP. 
String manipulations are error-prone and could even give rise to severe security vulnerabilities (e.g., cross-site scripting, aka XSS). 
One powerful method for identifying such bugs is \emph{symbolic execution}, which is possibly in combination with dynamic analysis. It analyses symbolic paths in a program by viewing them as constraints checked by constraint solvers. 
%
Symbolic execution of string manipulating programs has motivated the highly active research area of \emph{string constraint solving}, resulting in the development of numerous string solvers in the last decade, e.g.,
Z3seq~\cite{z3seq}, CVC4/5~\cite{cvc4}, Z3str/2/3/4~\cite{Z3-str,Z3-str2,Z3-str3,BerzishMurphy2021}, Z3str3RE~\cite{BD+23}, 
Z3-Trau~\cite{Z3-trau}, OSTRICH~\cite{CHL+19}, Slent~\cite{WC+18}, among many others.

@@

\medskip
\noindent 
\emph{Contribution.} In this work, we focus on the class of string constraints with regular membership and string-length function, where the counting operators may occur (called RECL for brevity). We make the following contributions: 
\vspace{-0.5mm}
\begin{itemize}
  \item We propose an automata-theoretical approach for solving RECL constraints. 
  Our main idea is to represent the counting operators by registers in cost-enriched finite automata (CEFA, see Section~\ref{sec:automaton} for the definition), instead of unfolding them explicitly. The satisfiability of RECL constraints is reduced to the nonemptiness problem of CEFA w.r.t. a linear integer arithmetic (LIA), which is then transformed to the satisfiability of existential LIA formula tackled by the off-the-shelf SMT solver.
  % Moreover, the string-length function can also be modeled by register in CEFA. Then the satisfiability of RECL constraints is reduced to the nonemptiness problem of CEFA with respect to a linear integer arithmetic (LIA) formula whose free variables are registers. According to the results from~\cite{atva2020}, an LIA formula can be computed to represent the potential values of registers in CEFA. 
  % Thus, the nonemptiness of CEFA w.r.t. LIA formulas can be reduced to the satisfiability of LIA formulas, which is then tackled by off-the-shelf SMT solvers.
  \item We propose simplification techniques to reduce the state space and the number of transitions in CEFA, to achieve better performance. It is beneficial since the sizes of LIA formulas depend much on the sizes of CEFA.
  \item Combined with the simplification techniques mentioned above, the register representation of regex-counting and string-length in CEFA entails an efficient algorithm for solving RECL constraints. We implement the algorithm on top of OSTRICH, resulting in a practical string solver called $\ostrichrecl$. 
  \item Finally, we utilize a collection of benchmark suites comprising 48,843 instances in total to evaluate the performance of our tool. The result shows that $\ostrichrecl$ solves the most instances among all the solvers, with comparable speed (See Table~\ref{tab:results_regcol}). In addition, we carry out some detailed experiments, including the evaluation of simplification techniques and the experiment of large-counting benchmarks. The result shows that the simplification techniques are effective in practice (See Table~\ref{tab:results_simp}), and $\ostrichrecl$ performs much better than other solvers on the large-counting benchmark (See Fig~\ref{fig:table_large_count}).
  % The experiment results show that $\ostrichrecl$ can solve 48,698 out of 48,843 instances, while Z3str3RE, the state-of-the-art solver on RECL constraints, solves only 47,607 instances (where the timeout period is set as 60 seconds). Moreover, the speed of $\ostrichrecl$ is comparable to that of Z3str3RE. In comparison, CVC5, Z3seq, Z3-Trau, and Z3str3 solve only 44,704, 46,426, 24,121, and 35,866 instances respectively. (See Table~\ref{tab:results_regcol}-\ref{tab:results_summary} for more details.)
\end{itemize}
\vspace{-0.5mm}
% We propose an automata-theoretical approach for solving RECL constraints. Our main idea is to represent the counting operators by registers in automata (as much as possible) instead of unfolding them explicitly. Our approach is built upon cost-enriched finite automata (CEFA), proposed in~\cite{atva2020} to solve string constraints involving the integer data type. Moreover, the string-length function can also be modeled by register in CEFA. Then the satisfiability of RECL constraints is reduced to the nonemptiness problem of CEFA with respect to a linear integer arithmetic (LIA) formula whose free variables are registers of CEFA.

@@

% Combined with the simplification techniques mentioned above, the register representation of regex-counting and string-length in CEFA entails an efficient algorithm for solving RECL constraints. We implement the algorithm on top of OSTRICH, resulting in a new version of OSTRICH called $\ostrichrecl$ for convenience. 
% Furthermore, we utilize a collection of benchmark suites comprising 48,843 RECL constraints in total to evaluate the performance of $\ostrichrecl$. It includes a widely-used benchmark suite for string constraints and benchmark suites comprising constraints generated from regular expressions in various open source regex libraries. 
% %
% %Note that in the benchmark suites, we only consider the regexes that contain explicit counting, that is, regexes that contain sub-expressions of the form $e^{\{n, m\}}$, and ignore those that contain only implicit counting operators, that is, $e^*$ or $e^+$, since the state-of-the-art string solvers suffer mainly from the explicit counting operators. 
% %
% The experiment results show that $\ostrichrecl$ can solve 48,698 out of 48,843 instances, while Z3str3RE, the state-of-the-art best solver on RECL constraints, solves only 47,607 instances (where the timeout period is set as 60 seconds). Moreover, the speed of $\ostrichrecl$ is comparable to that of Z3str3RE. In comparison, CVC5, Z3seq, Z3-Trau, and Z3str3 solve only 44,704, 46,426, 24,121, and 35,866 instances respectively. (See Table~\ref{tab:results_regcol}-\ref{tab:results_summary} for more details.)
